Predicting tree failure likelihood for utility risk mitigation via a novel convolutional neural network

### 2021-02-16 docs/manuscript/manuscript.tex

\usepackage{enumerate}

@@

\begin{document}
% ALL CAPS TITLE REQUIRED
\title{Predicting tree failure likelihood for utility risk mitigation via a novel convolutional neural network}
\author[1]{Author 1}
\author[2]{Author 2}
\author[2]{Author 3}
\author[2]{Author 4}
\author[2]{Author 5}

\affil[1]{First affiliation address, with corresponding author email. Email: author.one@umass.edu}
\affil[2]{Second affiliation address}

@@

\subsection{Pre-processing and data augmentation}
To achieve robustness in training, and given the relatively small number of training images, we randomly cropped each image on either axis to 3024 $\times$ 3024 pixels, generating five instances for each one. Thus, we increased the size of our training set from 505 to 2525 images. Further, we performed horizontal flipping with a 50\% probability on each of the generated images. For efficiency, we converted the images to grayscale and scaled the pixel values from 0 to 1. Finally, we downsampled the images to the following resolutions (pixels): $64 \times 64$, $128 \times 128$, $224 \times 224$ and $384\times 384$, creating a training set for each case. 
Random sets of images from each class across each of the four classification scenarios are shown in \autoref{fig:processed_images}.

@@

\subsection{Hyperparameter optimization}
We used the Hyperband approach \cite{li2018hyperband} to perform a grid search to find the optimal values of the following hyperparameters:
\begin{itemize}
    \item kernel size in first convolutional layer $ks$
    \item number of units in first densely connected layer $u_1$
    \item dropout rate applied to outputs of first dense layer $d_1$
    \item activation function for first dense layer $a_1$
    \item number of units in second densely connected layer $u_2$
    \item dropout rate applied to outputs of second dense layer $d_2$
    \item activation function for second dense layer $a_2$
    \item learning rate (of Adam optimizer) $\lambda$
\end{itemize}

\begin{table}[h!]\small
    \centering
    \begin{tabular}{l  c  p{1.75cm} p{1.75cm}  p{1.75cm} p{1.75cm}}\toprule
    \bf Scenario & \bf Hyperparameter & \multicolumn{3}{c}{\bf Resolution ($n\times n$ pixels)} \\
    &         & 64 & 128 & 224   \\\midrule
    \multirow{7}{*}{\texttt{Pr\_Po\_Im}} 
        & $ks$    & 7 &  &     \\
        & $u_1$   & 128 &  &     \\
        & $d_1$   & 0.25 &  &     \\
        & $a_1$   & ReLu  &  &      \\
        & $u_2$   & 320 &  &    \\
        & $d_2$   & 0.30 &  &     \\
        & $a_2$   & ReLu &  &     \\
        & $\lambda$  & $1.76\cdot 10^{-4}$ &  &      \\ \midrule
    \multirow{7}{*}{\texttt{Pr\_Im}} 
        & $ks$    & 7 &  &      \\
        & $u_1$   & 128 &  &     \\
        & $d_1$   & 0.25 &  &      \\
        & $a_1$   & ReLu  &  &      \\
        & $u_2$   & 320 &  &      \\
        & $d_2$   & 0.30 &  &     \\
        & $a_2$   & ReLu &  &      \\
        & $\lambda$  & $1.76\cdot 10^{-4}$ &  &      \\ \midrule
    \multirow{7}{*}{\texttt{PrPo\_Im}} 
        & $ks$    & 7 &  &    \\
        & $u_1$   & 128 &  &     \\
        & $d_1$   & 0.25 &  &      \\
        & $a_1$   & ReLu  &  &     \\
        & $u_2$   & 320 &  &     \\
        & $d_2$   & 0.30 &  &     \\
        & $a_2$   & ReLu &  &      \\
        & $\lambda$  & $1.76\cdot 10^{-4}$ &  &     \\ \midrule
    \multirow{7}{*}{\texttt{Pr\_PoIm}}  
        & $ks$    & 7 &  &      \\
        & $u_1$   & 128 &  &      \\
        & $d_1$   & 0.25 &  &      \\
        & $a_1$   & ReLu  &  &      \\
        & $u_2$   & 320 &  &      \\
        & $d_2$   & 0.30 &  &     \\
        & $a_2$   & ReLu &  &      \\
        & $\lambda$  & $1.76\cdot 10^{-4}$ &  &  &   \\ \bottomrule
    \end{tabular}
    \caption{Optimal hyperparameters found using the Hyperband search algorithm for all classification scenario and input image resolution combinations.}
    \label{tab:opt-hyp}
\end{table}
\section{Results}

@@

\begin{figure}
    \centering
    \includegraphics{}
    \caption{Caption}
    \label{fig:my_label}
\end{figure}

>>>>

\usepackage{enumerate}
\usepackage{siunitx}

@@

\begin{document}
% ALL CAPS TITLE REQUIRED
\title{Predicting tree failure likelihood for utility risk mitigation via a novel convolutional neural network}
\author[1]{Author 1}
\author[1]{Author 2}
\author[2]{Author 3}
\author[2]{Author 4}
\author[2]{Author 5}

\affil[1]{Department of Civil and Environmental Engineering, University of Massachusetts Amherst, MA 01003, USA} %. Email: author.one@umass.edu}
\affil[2]{Department of Ecological Conservation, University of Massachusetts Amherst, MA 01003, USA}

@@

\subsection{Pre-processing and data augmentation}
To achieve robustness in training, and given the relatively small number of training images, we randomly cropped each image on either axis to 3024 $\times$ 3024 pixels, generating five instances for each one. Thus, we increased the size of our training set from 505 to 2525 images. Further, we performed horizontal flipping with a 50\% probability on each of the generated images. For efficiency, we converted the images to grayscale and scaled the pixel values from 0 to 1. Finally, we downsampled the images to the following resolutions (pixels): $64 \times 64$, $128 \times 128$ and $224 \times 224$, creating a training set for each case. 
Random sets of images from each class across each of the four classification scenarios are shown in \autoref{fig:processed_images}.

@@

\subsection{Hyperparameter optimization}
We used the Hyperband approach \cite{li2018hyperband} to perform a guided grid search to find the optimal hyperparameter values across each scenario and input resolution instance. Thus, we performed 12 searches on the following hyperparameters in our CNN:
\begin{itemize}
    \item kernel size in first convolutional layer $ks$
    \item number of units in first densely connected layer $u_1$
    \item dropout rate applied to outputs of first dense layer $d_1$
    \item activation function for first dense layer $a_1$
    \item number of units in second densely connected layer $u_2$
    \item dropout rate applied to outputs of second dense layer $d_2$
    \item activation function for second dense layer $a_2$
    \item learning rate (of Adam optimizer) $\lambda$
\end{itemize}
Each search was conducted using 90 trials of unique hyperparameter combinations.

\begin{table}[h!]\small
    \centering
    \sisetup{round-mode=places,
        round-precision=2,
        scientific-notation=fixed,
        exponent-product = \cdot,
        table-omit-exponent,
        fixed-exponent=-4
        }
 % Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
 \begin{tabular}{@{}llr r@{}}
\toprule
& & \multicolumn{2}{c}{\bf Resolution}\\
\textbf{Scenario}   & \textbf{Hyperparameters} & 64                & 128               \\ \midrule
PrPo\_Im   & activation\_1   & relu              & tanh              \\
   & activation\_2   & relu              & relu              \\
   & dropout\_1      & 0.1               & 0.25              \\
   & dropout\_2      & 0.3               & 0.35              \\
   & kernel\_size    & 7                 & 5                 \\
   & learning\_rate  & \num{0.000102956216062} & \num{0.000109137476524} \\
   & units\_1        & 480               & 384               \\
   & units\_2        & 448               & 256               \\
Pr\_Im     & activation\_1   & relu              & relu              \\
      & activation\_2   & relu              & relu              \\
      & dropout\_1      & 0.1               & 0.1               \\
     & dropout\_2      & 0.3               & 0.3               \\
      & kernel\_size    & 7                 & 7                 \\
      & learning\_rate  & \num{0.000102956216062} & \num{0.000102956216062} \\
      & units\_1        & 480               & 480               \\
      & units\_2        & 448               & 448               \\
Pr\_PoIm   & activation\_1   & relu              & relu              \\
   & activation\_2   & tanh              & relu              \\
   & dropout\_1      & 0.25              & 0.1               \\
   & dropout\_2      & 0.3               & 0.3               \\
   & kernel\_size    & 7                 & 7                 \\
   & learning\_rate  & \num{0.000175822328104} & \num{0.000102956216062} \\
   & units\_1        & 128               & 480               \\
   & units\_2        & 320               & 448               \\
Pr\_Po\_Im & activation\_1   & relu              & tanh              \\
  & activation\_2   & tanh              & relu              \\
  & dropout\_1      & 0.25              & 0.25              \\
  & dropout\_2      & 0.3               & 0.35              \\
  & kernel\_size    & 7                 & 5                 \\
  & learning\_rate  & \num{0.000175822328104} & \num{0.000109137476524} \\
  & units\_1        & 128               & 384               \\
  & units\_2        & 320               & 256              \\\bottomrule
\end{tabular}
     % \begin{tabular}{l  c  p{1.75cm} p{1.75cm}  p{1.75cm} p{1.75cm}}\toprule
    % \bf Scenario & \bf Hyperparameter & \multicolumn{3}{c}{\bf Resolution ($n\times n$ pixels)} \\
    % &         & 64 & 128 & 224   \\\midrule
    % \multirow{7}{*}{\texttt{Pr\_Po\_Im}} 
    %     & $ks$    & 7 &  &     \\
    %     & $u_1$   & 128 &  &     \\
    %     & $d_1$   & 0.25 &  &     \\
    %     & $a_1$   & ReLu  &  &      \\
    %     & $u_2$   & 320 &  &    \\
    %     & $d_2$   & 0.30 &  &     \\
    %     & $a_2$   & ReLu &  &     \\
    %     & $\lambda$  & $1.76\cdot 10^{-4}$ &  &      \\ \midrule
    % \multirow{7}{*}{\texttt{Pr\_Im}} 
    %     & $ks$    & 7 &  &      \\
    %     & $u_1$   & 128 &  &     \\
    %     & $d_1$   & 0.25 &  &      \\
    %     & $a_1$   & ReLu  &  &      \\
    %     & $u_2$   & 320 &  &      \\
    %     & $d_2$   & 0.30 &  &     \\
    %     & $a_2$   & ReLu &  &      \\
    %     & $\lambda$  & $1.76\cdot 10^{-4}$ &  &      \\ \midrule
    % \multirow{7}{*}{\texttt{PrPo\_Im}} 
    %     & $ks$    & 7 &  &    \\
    %     & $u_1$   & 128 &  &     \\
    %     & $d_1$   & 0.25 &  &      \\
    %     & $a_1$   & ReLu  &  &     \\
    %     & $u_2$   & 320 &  &     \\
    %     & $d_2$   & 0.30 &  &     \\
    %     & $a_2$   & ReLu &  &      \\
    %     & $\lambda$  & $1.76\cdot 10^{-4}$ &  &     \\ \midrule
    % \multirow{7}{*}{\texttt{Pr\_PoIm}}  
    %     & $ks$    & 7 &  &      \\
    %     & $u_1$   & 128 &  &      \\
    %     & $d_1$   & 0.25 &  &      \\
    %     & $a_1$   & ReLu  &  &      \\
    %     & $u_2$   & 320 &  &      \\
    %     & $d_2$   & 0.30 &  &     \\
    %     & $a_2$   & ReLu &  &      \\
    %     & $\lambda$  & $1.76\cdot 10^{-4}$ &  &  &   \\ \bottomrule
    % \end{tabular}
    \caption{Optimal hyperparameters found using the Hyperband search algorithm for all classification scenario and input image resolution combinations.}
    \label{tab:opt-hyp}
\end{table}
\section{Results}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/scenario_resolution_performance.png}
    \caption{Performance metrics for each scenario and input resolution instance. Average metrics and confidence intervals (trials, $n=10$) are shown for 10 epochs in each case.}
    \label{fig:sens-overall}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/scenario_resolution_validation_accuracy.png}
    \caption{Mean validation accuracy values with error bars for each scenario and resolution combination.}
    \label{fig:sens-val-acc}
\end{figure}
